Rapid, (semi-?)Supervised Generation of Novel Percussive Sounds With A Tractable Virtual Synthesizer

### 2020-01-02 NIME_PAPER/paper.tex
section 2

\subsection{The Ear}
We defined an ear as any method of evaluating a piece of audio, capable of "listening" to a piece of audio and giving it a score (or a list of scores) based on how well it satisfies certain criteria. Since in this work we are mainly focused on percussive generation, what we require from the ear is to give us probabilities of an audio sample belonging to various sound (drum and none-drum) categories. As our synthesizer outputs are deterministic for all programs, allowing us to associate the categorical probabilities of each sound with the program that generated it; navigating our synthesis towards parameters which give us the desired sonic output.\\
% \begin{figure}[H]
% \centering
% \includegraphics[width=\textwidth]{images/SSS_ear.png}
% \caption{Evaluation Using the Ear}
% \label{fig:SSS generator}
% \end{figure}
%  The most basic representation of a piece of digital audio is via list of numbers and a sampling-rate. For simplicity, we fix the sampling rate to 41,000 per second across all components of this project. An array representation of audio is the representation of the audio within the time domain. More commonly, the frequency domain is used for MIR tasks.
We have implemented several methods of feature extraction and successfully separated different drum categories in interesting ways. 
\todo[inline, color=green!30]{is this the right place to show off some t-sne graphs of how various feature extraction methods seperated drum categories?}
A common pitfall of most feature extraction methods implemented is that they do not consider the temporal trends within the audio, making them highly susceptible to errors when evaluating audio that should not belong to any of the categories.\\
Analysis of the frequency content of a piece of audio using windowing functions is a common MIR procedure. Care must be taking such that information critical to percussion categorization such as envelope and temporal changes are not lost in this process. For example, we do not want a reversed kick sample to be labeled as a kick, or even a drum, even if it has the correct frequency content. There are two interconnected temporal trends to consider: 
\begin{itemize}
    \item Change in amplitude within each frequency bin. Harder to measure and utilize and possibly not as important for our purposes, considering the short length of our samples.
    \item Overall amplitude change within a sample. Closely linked to ADSR\\. 
\end{itemize}{}
Temporal trends may be deduced from a spectrum representation of sound. A convolution net trained on our drum dataset, combined with an algorithm ADSR extraction algorithm led to an ear which can reliably reject junk data and samples with inappropriate wave-shapes. \todo[inline,color=green!30]{This paragraph is lying, I haven't done the ADSR algorithm part yet. Highest priority after writing this paper.}
\begin{figure}[H]
\centering
\includegraphics[width=0.8\linewidth]{images/cnn_ear.png}
\caption{Model working well on a test set}
\label{fig:ok ear}
\end{figure}
\todo[inline,color=green!30]{ will add f-measure calculations and other confidence inducing graphs (for example I have a grid of spectrums with the ear's guess vs the actual label under each spectrum, I can colorize the grid depending on whether the guess was correct or not). Code in the notebooks. After updating ear, I will make them look presentable and add them.}
\subsection{How Can The Synth and Ear Interact?}
With our Synth and Ear in place, we can quickly generate programs for our Synth, create the corresponding audio sample and get the scores for the audio sample. If the ear returns a vector of categorical probabilities, we can assign categorical ranks to the audio sample. For example, the category with the highest probability will have the rank of 1, second highest would rank 2nd, and the lowest probability will have the rank of n (assuming n categories).\textbf{
But how can we use this to create percussion sounds?}\\\\
We utilize our tools in three different methods of generation, with their specific imlpementation discussed in section \ref{gens}.
\begin {enumerate} [label=(\roman*)]
\item Random Generation: We randomly create a drum programs until a sound belonging to our category is produced. We hypothesize that due to the speed of our experiments, it is a viable approach.
\item Multivariate Distribution Sampling: We attempt to learn the joint distribution of our parameters in order to maximize generation of a certain drum category. 
\item Genetic Search: We treat our parameters as genes and attempt to use evolutionary search algorithms to find programs that the ear deems most fit.\\
\end {enumerate} 
\todo[inline,color=blue!30]{As seen in Figure \ref{fig:rank portions}, there is a bias in random generation towards making higher ranking hats and kicks and lower ranking shakers, claps. Is this bias because of Synth structure or a problem with the ear? Will analyse further after updating ear}
\begin{figure}[h!]
\centering
\includegraphics[width=1\linewidth]{images/random_ranks.png}
\caption{Figure showing proportions of categorical ranks for around 50,000 randomly generated sounds. Only showing ranks 1-5. About 90\% received the rank of 1 (blue portion of the bar) for the "synth" category, meaning that the ear correctly gave the highest probability to the sound being from a synth combination. We are interested in the small percentages of random generations that trick the ear due to coincidentally being similar to actual drums (i.e those who received a rank of 1) }
\todo[inline, color=blue!30]{When looking for the best sounds of each category, could use ranks 2 and 3 etc. as well, however:
\unexpanded{\unexpanded{
\begin{itemize}
    \item We must figure out how to penalize their parameters
    \item The task requires even more confidence in ear,
\end{itemize} 
}}}

\label{fig:rank portions}
\end{figure}
\section{Novel Generations}
\label{gens}
\subsection{Goal}
 Given a category, our aim is to use random generations to guide an algorithm towards novel generation of drums for that category.

\subsection{Guided Generation Attempts}

>>>>

\subsection{The Ear}
We defined an ear as any method of evaluating a piece of audio, capable of "listening" to a piece of audio and giving it a score (or a list of scores) based on how well it satisfies certain criteria. Since in this work we are mainly focused on percussive generation, what we require from the ear is to give us probabilities of an audio sample belonging to various sound (drum and none-drum) categories. As our synthesizer outputs are deterministic for all programs, allowing us to associate the categorical probabilities of each sound with the program that generated it; navigating our synthesis towards parameters which give us the desired sonic output.\\

%  The most basic representation of a piece of digital audio is via list of numbers and a sampling-rate. For simplicity, we fix the sampling rate to 41,000 per second across all components of this project. An array representation of audio is the representation of the audio within the time domain. More commonly, the frequency domain is used for MIR tasks.
We have implemented several methods of feature extraction and successfully separated different drum categories in interesting ways. 
\todo[inline, color=green!30]{is this the right place to show off some t-sne graphs of how various feature extraction methods seperated drum categories?}
A common pitfall of most feature extraction methods implemented is that they do not consider the temporal trends within the audio, making them highly susceptible to errors when evaluating audio that should not belong to any of the categories.\\
Analysis of the frequency content of a piece of audio using windowing functions is a common MIR procedure. Care must be taking such that information critical to percussion categorization such as envelope and temporal changes are not lost in this process. For example, we do not want a reversed kick sample to be labeled as a kick, or even a drum, even if it has the correct frequency content. There are two interconnected temporal trends to consider: 
\begin{itemize}
    \item Change in amplitude within each frequency bin. Harder to measure and utilize and possibly not as important for our purposes, considering the short length of our samples.
    \item Overall amplitude change within a sample. Closely linked to ADSR\\. 
\end{itemize}{}
Temporal trends may be deduced from a spectrum representation of sound. A convolution net trained on our drum dataset, combined with an algorithm ADSR extraction algorithm led to an ear which can reliably reject junk data and samples with inappropriate wave-shapes. \todo[inline,color=green!30]{This paragraph is lying, I haven't done the ADSR algorithm part yet. Highest priority after writing this paper.}
\begin{figure}[H]
\centering
\includegraphics[width=0.8\linewidth]{images/cnn_ear.png}
\caption{Model working well on a test set}
\label{fig:ok ear}
\end{figure}
\todo[inline,color=green!30]{ will add f-measure calculations and other confidence inducing graphs (for example I have a grid of spectrums with the ear's guess vs the actual label under each spectrum, I can colorize the grid depending on whether the guess was correct or not). Code in the notebooks. After updating ear, I will make them look presentable and add them.}
%
% How Can The Synth and Ear Interact?
%
\subsection{How Can The Synth and Ear Interact?}

\begin{figure*}[htbp]
	\centering
		\includegraphics[width=0.7\textwidth]{images/SSS_gen.png}
	\caption{\colorbox{green!=40}{ This graph is bad in many ways.} but would it be worthwhile to have a \textbf{good} flowchart of 
	the generation pipeline in the paper? }
	\label{fig:BlockDiagram2}
\end{figure*}
\label{fig:generation}

With our Synth and Ear in place, we can quickly generate programs for our Synth, create the corresponding audio sample and get the scores for the audio sample. If the ear returns a vector of categorical probabilities, we can assign categorical ranks to the audio sample. For example, the category with the highest probability will have the rank of 1, second highest would rank 2nd, and the lowest probability will have the rank of n (assuming n categories).\textbf{
But how can we use this to create percussion sounds?}\\\\
We utilize our tools in three different methods of generation, with their specific imlpementation discussed in section \ref{gens}.
\begin {enumerate} [label=(\roman*)]
\item Random Generation: We randomly create a drum programs until a sound belonging to our category is produced. We hypothesize that due to the speed of our experiments, it is a viable approach.
\item Multivariate Distribution Sampling: We attempt to learn the joint distribution of our parameters in order to maximize generation of a certain drum category. 
\item Genetic Search: We treat our parameters as genes and attempt to use evolutionary search algorithms to find programs that the ear deems most fit.\\
\end {enumerate} 
\todo[inline,color=blue!30]{As seen in Figure \ref{fig:rank portions}, there is a bias in random generation towards making higher ranking hats and kicks and lower ranking shakers, claps. Is this bias because of Synth structure or a problem with the ear? Will analyse further after updating ear}
\begin{figure}[h!]
\centering
\includegraphics[width=1\linewidth]{images/random_ranks.png}
\caption{\colorbox{green!=40}{will update and make labels bigger} Showing proportions of categorical ranks for around 50,000 randomly generated sounds. Only showing ranks 1-5. About 90\% received the rank of 1 (blue portion of the bar) for the "synth" category, meaning that the ear correctly gave the highest probability to the sound being from a synth combination. We are interested in the small percentages of random generations that trick the ear due to coincidentally being similar to actual drums (i.e those who received a rank of 1) }
\todo[inline, color=blue!30]{When looking for the best sounds of each category, could use ranks 2 and 3 etc. as well, however:
\unexpanded{\unexpanded{
\begin{itemize}
    \item We must figure out how to penalize their parameters
    \item The task requires even more confidence in ear,
\end{itemize} 
}}}

\label{fig:rank portions}
\end{figure}
\section{Novel Generations}
\label{gens}
\subsection{The Goal}
 Given a category, our aim is to use random generations to guide an algorithm towards novel generation of drums for that category.

\subsection{Methodologies}
\subsubsection{Random Generation}
